---
title: "Clinical Data Validation in R"
subtitle: "A guide on how to do clinical data validation using some R packages."
author: "Jeremy Selva"
---

{{< include ./_extensions/r-wasm/live/_knitr.qmd >}}
{{< include ./_extensions/r-wasm/live/_gradethis.qmd >}}

# Preface {.unnumbered}

This website provides structured examples of ways to do clinical data validation using some R packages such as [`pointblank`](https://rstudio.github.io/pointblank/){target="_blank"}.

## Who is this for?

This is not a beginner's introduction to R or data programming in general, though you do not need to be an expert to understand its content If you have some familiarity with tabular clinical data, most of the examples should make sense.

## R Packages Used

Here are the R packages used in this website.

```{r}
#| label: R packages load
#| message: false
#| echo: false

library(sessioninfo)
library(fs)
# library(quarto)
library(tibble)
library(dplyr)
library(tidyr)
library(purrr)
# library(forcats)
# library(glue)
# library(readxl)
library(checkmate)
library(reactable)
library(pointblank)

```

```{r}
#| label: R packages used
#| message: false
#| echo: false

r_package_table <- sessioninfo::package_info()
rownames(r_package_table) <- NULL

r_package_table <- r_package_table |>
  tibble::as_tibble() |>
  dplyr::mutate(
    version = ifelse(
      is.na(r_package_table$loadedversion),
      r_package_table$ondiskversion,
      r_package_table$loadedversion)
  ) |>
  dplyr::filter(.data[["attached"]] == TRUE) |>
  dplyr::select(
    dplyr::any_of(
      c("package", "version", "date", "source")
    )
  )

reactable::reactable(
  r_package_table,
  defaultPageSize = 5
)

```

## R Platform Information

Here are the R platform environment used in this analysis.

```{r}
#| label: R platform used
#| echo: false

get_knitr_version <- function() {

  knitr_info <- "NA"

  r_package_table <- sessioninfo::package_info(
    pkgs = c("installed")
  ) |>
    dplyr::filter(.data[["package"]] == "knitr")

  if (nrow(r_package_table) == 1) {

    knitr_version <- r_package_table$loadedversion[1]
    knitr_source <- r_package_table$source[1]

    knitr_info <- paste0(
      knitr_version, " from ",
      knitr_source)
  }

  return(knitr_info)

}

r_platform_table <- suppressWarnings(sessioninfo::platform_info())
r_platform_table[["knitr"]] <- get_knitr_version()[1]

r_platform_table <- data.frame(
  setting = names(r_platform_table),
  value = unlist(r_platform_table,
                 use.names = FALSE),
  stringsAsFactors = FALSE
)


reactable::reactable(
  r_platform_table,
  defaultPageSize = 5
)
```

## Credit

[Photo](https://stocksnap.io/photo/notepad-notebook-RMLFV5PVU9){target="_blank"} by [Dariusz Sankowski](https://stocksnap.io/author/16700){target="_blank"} on [StockSnap](https://stocksnap.io).

